---
title: "case_study_bellabeat"
author: "LJ"
date: "2023-12-27"
output: html_document
---
# Capstone Project: BellaBeat's Smart Device Data Analysis

# Table of Content
* Summary
* Data Analysis Roadmap
  + Ask
  + Prepare
  + Process
  + Analyze
  + Share
  + Act
* Conclusion

## Summary

### About The Company
Bellabeat, founded by Urška Sršen and Sando Mur, is a high-tech company specializing in health-focused smart products. The company develops beautifully designed technology aimed at informing and inspiring women worldwide. Bellabeat's products collect data on activity, sleep, stress, and reproductive health, empowering women with knowledge about their well-being and habits.

### Business Tasks
Bellabeat offers several products: Bellabeat app, Leaf (wellness tracker), Time (wellness watch), Spring (water bottle), and membership. The project's goal is to focus on one of Bellabeat’s products and analyze smart device data to gain insight into how consumers are using their smart devices. Specifically, the project will include analysis of Leaf and Time.

### Stakeholders
* Urška Sršen: Bellabeat’s cofounder and Chief Creative Officer
* Sando Mur: Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team

## Data Analysis Roadmap
### Ask
In order to address the major business tasks, the following questions regarding the smart devices' usage need to the answered:
* What are some trends in smart device usage?
* How could these trends apply to Bellabeat customers?
* How could these trends help influence Bellabeat marketing strategy?

The datasets are obtained from [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit/data), author [Möbius](https://www.kaggle.com/arashnic). This Kaggle data set
contains personal fitness tracker from thirty fitbit users. Thirty eligible Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring. It includes information about daily activity, steps, and heart rate that can be used to explore users’ habits.


### Prepare
* ROCCC Check
  ROCCC stands for Reliable, Original, Comprehensive, Current, and Cited. The dataset is an open-source, third-party one. This dataset generated by respondents to a distributed survey via Amazon Mechanical Turk between 03.12.2016-05.12.2016: thirty eligible Fitbit users consented to the submission of personal tracker data. Therefore, the dataset reaches a moderate degree of quality in all aspects and it's slightly better in Reliability and Citedness.
  
* Data Structure and Data Type
  The datasets that will be analyzed (daily and hourly data) contain long data. The data types are mainly numbers and data/time objects.

Below is the code to get a glimpse of the datasets used:
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
```

```{r}
# load datasets
# daily & hourly
daily_activity <- read.csv('dailyActivity_merged.csv')
daily_calories <- read.csv('dailyCalories_merged.csv')
daily_intensities <- read.csv('dailyIntensities_merged.csv')
daily_steps <- read.csv('dailySteps_merged.csv')
hourly_calories <- read.csv('hourlyCalories_merged.csv')
hourly_intensities <- read.csv('hourlyIntensities_merged.csv')
hourly_steps <- read.csv('hourlySteps_merged.csv')
```

```{r}
# check datasets
glimpse(daily_activity)
glimpse(daily_calories)
glimpse(daily_intensities)
glimpse(daily_steps)
glimpse(hourly_calories)
glimpse(hourly_intensities)
glimpse(hourly_steps)
```

From the above check on the contents of the datasets, we now know the size of each dataset and which kind of data is included in the datasets. We also know where in the datasets that we need to clean in the next stage. For instance, in the `hourly_` datasets, we need to separate date from hours to make sure that each row corresponds to an hour in a date.